### 2015-07-10 thesis/thesis/abstract.tex
changes

 \textbf{This is not real abstract} In the last few years, communities of services have been studied in a certain numbers of proposals as virtual pockets of similar expertise. The motivation is to provide these services with high chance of discovery through better visibility, and to enhance their capabilities when it comes to provide requested functionalities. There are a number of proposed mechanisms and models on aggregating web services and making them cooperate within their communities. However, forming optimal and stable communities as coalitions to maximize individual and group efficiency and income for all the involved parties has not been addressed yet. Also, in the proposed frameworks of these communities, a common assumption is that residing services, which are supposed to be autonomous and intelligent, are competing over received requests. However, those services can also exhibit cooperative behaviors, for instance in terms of substituting each other. Deciding to compete or cooperate inside communities is a problem yet to be investigated.

>>>>

In the last few years, communities of services have been studied in a certain numbers of proposals as virtual pockets of similar expertise. The motivation is to provide these services with high chance of discovery through better visibility, and to enhance their capabilities when it comes to provide requested functionalities. There are a number of proposed mechanisms and models on aggregating web services and making them cooperate within their communities. However, forming optimal and stable communities as coalitions to maximize individual and group efficiency and income for all the involved parties has not been addressed yet. Also, in the proposed frameworks of these communities, a common assumption is that residing services, which are supposed to be autonomous and intelligent, are competing over received requests. However, those services can also exhibit cooperative behaviors, for instance in terms of substituting each other. Deciding to compete or cooperate inside communities is a problem yet to be investigated.

### 2015-07-10 thesis/thesis/chap2/chap2.tex
changes

\chapter{Background}\label{cha:background}
In this chapter, we briefly review web services, then we introduce
the concept of communities of web services, their architecture and
applications and the benefits of forming communities. Thereafter,
we discuss the cooperative game theory concepts used throughout
the proposal. Finally, we discuss relevant related work on web
service communities and games in the literature of service
oriented computing.

@@

We are not going to delve into engineering details of online web
service implementation and its protocols in this proposal. We are
interested in web services from their business model perspective.
Service providers usually charge end users for services they
provide. For example, Google has listed their pricing and plans
for wide range of services they provide on their web service
console page\footnote{https://code.google.com/apis/console}.

@@

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed
mechanisms to help users and services maximize their gain. A
two-player non-cooperative game between web services and community
master was introduced in
\cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this game-theoretic
model, the strategies available to a web service when facing a new
community are requesting to join the community, accepting the
master's invitation to join the community, or refusing the
invitation to join. The set of strategies for communities are
inviting the web service or refusing the web service's join
request. Based on their capacity, market share and reputation, the
two players have different sets of utilities over the strategy
profiles of the game. The main limits of this game model are: 1)
its consideration of only three quality parameters, while the
other factors are simply ignored; and 2) the non-consideration of
the web services already residing within the community. The game
is only between the community master and the new web service, and
the inputs from all the other members and their influence on the
master's decision are simply ignored. The consideration of those
inputs and this influence factor is a significant issue as
existing web services can lose utility or payoff because of the
new member, which can result in an unhealthy and unstable group.
The problem comes from the fact that the existing members should
collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our paper.

>>>>

\chapter{Background}\label{cha:background}
In this chapter, we briefly review web services, then we introduce
the concept of communities of web services, their architecture and
applications and the benefits of forming communities. Thereafter,
we discuss the cooperative game theory concepts used throughout
the thesis. Finally, we discuss relevant related work on web
service communities and games in the literature of service
oriented computing.

@@

We are not going to delve into engineering details of online web
service implementation and its protocols in this thesis. We are
interested in web services from their business model perspective.
Service providers usually charge end users for services they
provide. For example, Google has listed their pricing and plans
for wide range of services they provide on their web service
console page\footnote{https://code.google.com/apis/console}.

@@

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed
mechanisms to help users and services maximize their gain. A
two-player non-cooperative game between web services and community
master was introduced in
\cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this game-theoretic
model, the strategies available to a web service when facing a new
community are requesting to join the community, accepting the
master's invitation to join the community, or refusing the
invitation to join. The set of strategies for communities are
inviting the web service or refusing the web service's join
request. Based on their capacity, market share and reputation, the
two players have different sets of utilities over the strategy
profiles of the game. The main limits of this game model are: 1)
its consideration of only three quality parameters, while the
other factors are simply ignored; and 2) the non-consideration of
the web services already residing within the community. The game
is only between the community master and the new web service, and
the inputs from all the other members and their influence on the
master's decision are simply ignored. The consideration of those
inputs and this influence factor is a significant issue as
existing web services can lose utility or payoff because of the
new member, which can result in an unhealthy and unstable group.
The problem comes from the fact that the existing members should
collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our research work.

### 2015-07-10 thesis/thesis/chap3/chap3.tex
changes

In this section, we discuss the parameters and preliminary
concepts that we use in the rest of the proposal.

@@

\emph{1) Web services} are rational entities that aim to maximize
their utilities by providing high quality services to end users.
They aim to maximize their individual income by receiving enough
requests from end users. In order to increase their revenue, web
services seek for more tasks if they have the capacity and
throughput to do so. Web services can join communities to have
better efficiency by collaborating with others, to have access to
higher market share, and to have opportunity of receiving a bigger
task pool from end users. Throughout this proposal, in our
equations, we refer to web services as $ws$ and to the set of web
services hosted by a given community as $C$. To simplify the
notation, sometimes we simply write $ws$ instead of $ws \in C$ to
go through the elements $ws$ of the set $C$.

>>>>

In this section, we discuss the parameters and preliminary
concepts that we use in the rest of the chapter.

@@

\emph{1) Web services} are rational entities that aim to maximize
their utilities by providing high quality services to end users.
They aim to maximize their individual income by receiving enough
requests from end users. In order to increase their revenue, web
services seek for more tasks if they have the capacity and
throughput to do so. Web services can join communities to have
better efficiency by collaborating with others, to have access to
higher market share, and to have opportunity of receiving a bigger
task pool from end users. Throughout this thesis, in our
equations, we refer to web services as $ws$ and to the set of web
services hosted by a given community as $C$. To simplify the
notation, sometimes we simply write $ws$ instead of $ws \in C$ to
go through the elements $ws$ of the set $C$.

### 2015-07-10 thesis/thesis/chap4/chap4.tex
changes

\textbf{Contributions.} In this paper, we introduce DDM, a Distributed Decision Making model for community formation that regulates web service agents’ decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. Unlike existing work on community formation, our decision model is extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and how to select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree, which is initially created from the past data, reflects a comprehensive vision about agents' attitudes in terms of their action selection based on their past experiences. Moreover, the tree is getting continuously updated based on both new received feedback and the outcome of chosen actions. This continuous update makes the approach adapted to any change in the environment.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems.
The decision model provides web services with enough information which helps those services efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in their decision making and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. The proposed model can be seen as a recommneder system that suggests beneficial actions for both communities and single services. Communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact both their short-term and long-term goals; therefore, opting for the best decision available.

@@

We are not going to delve into the engineering details of online web service implementation and its protocols in this paper. We are interested in web services from a business model perspective. Service providers usually charge end users for services they provide. For example, Google has listed pricing and plans for a wide range of services they provide on their web service console page\footnote{$https://code.google.com/apis/console$}.

@@

In this section, we discuss the parameters that we use in the rest of the paper. Then, we present the task distribution and revenue model of our distributed web services communities.

@@

With a group of web services having identical or similar functionalities, QoS metrics provide nonfunctional characteristics for optimal candidate selection. Web services quality metrics have been studied and analyzed in various proposals, for instance in in \cite{Ardagna:2007:ASC:1263152.1263531,Menasce:2002:QIW:613357.613758,10.1109/ISSRE.2011.17}. In this paper, we adopt the most representative QoS properties of those services that highly influence their utility.
%We refer to a typical web service as ws_{i}.

@@

The closest related work \cite{10.1109/TSC.2012.12,
DBLP:conf/IEEEscc/LimTMB12, DBLP:conf/IEEEscc/KhosravifarABT11} and our
previous work \cite{journal-community-formation} regarding the community formation problem have considered a centralized approach where a community manager has complete information of all the web services and their quality
metric and parameters. Those proposals run complex algorithms through all the space of
solutions in order to find the optimal answer. However, in this paper, we
have considered an unexplored and more realistic situation where information is incomplete and a decision
profile is generated based on a smaller set of web services. Our solution helps
communities and web services select actions that lead towards
maximizing their utilities. Therefore, considering the different settings,
we cannot experimentally compare our work with the mentioned related work.

@@

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed
mechanisms to help users and services maximize their gain. A
two-player non-cooperative game between web services and community
master was introduced in
\cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this game-theoretic
model, the strategies available to a web service when facing a new
community are requesting to join the community, accepting the
master's invitation to join the community, or refusing the
invitation to join. The set of strategies for communities are
inviting the web service or refusing the web service's join
request. Based on their capacity, market share and reputation, the
two players have different sets of utilities over the strategy
profiles of the game. The main limits of this game model are: 1)
its consideration of only three quality parameters, while the
other factors are simply ignored; and 2) the non-consideration of
the web services already residing within the community. The game
is only between the community master and the new web service, and
the inputs from all the other members and their influence on the
master's decision are simply ignored. The consideration of those
inputs and this influence factor is a significant issue as
existing web services can lose utility or payoff because of the
new member, which can result in an unhealthy and unstable group.
The problem comes from the fact that the existing members should
collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our paper.

@@

In \cite{managing-hela-jalel}, Limam and Akaichi have proposed web
service communities with centralized access across distributed web
services. They have proposed a framework for web service
management, query resolution among communities and a query caching
mechanism executed by the manager to improve the performance of
query resolution process among many distributed communities. The
key idea is to cache previous computed results for answering
future queries. Maamar et al. initially in \cite{conf/webist/MaamarLBTS07} and
then comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}
proposed an architecture utilizing \emph{Contract-Net} protocol
for engineering task distribution within communities of web
services. The protocol is centrally executed by the community
manager. This architecture has been further extended in
\cite{CSTintercommunity, conf/IEEEscc/BenharrefSBB11,
conf/IEEEscc/KhosravifarBMMT10, conf/aina/LimTM11}. Two types of
roles have been distinguished for community members: masters and
slaves. Master web services are community managers that lead
communities and are responsible for membership management. They
can invite and convince slave web services to join the community,
and attract new slave web services to their communities by
awarding them better payoff. Moreover, they can eject some slave
members from the community to improve its overall reputation if
these members are misbehaving or cannot provide the promised QoS.
In \cite{Medjahed05adynamic}, Medjahed and Bouguettaya have
developed a community as a ``cluster'' that groups Web services
based on a specific area of interest. All web services in a given
community share the same functionality. These communities are
created by \emph{third party community providers} which use the
\emph{community ontology} as a template and define a set of
operations that all web services within a community should
provide. Using semantic analysis on web service operations, web
services either find and join a community with similar
functionality or create a new operation description for a new
community. The authors have described the concept of
\emph{community agents} associated to \emph{community providers}.
A community agent is responsible, among other things, of the
registration of services with the community. An example of a
community that provides health care services to senior citizens
has been used. In this example, a governmental entity is needed to
check the health care standards used by the members before
authorizing them to be part of the community. Such a central
entity is represented by the community agent. Thus, community
agents are playing the role of community managers. In a close work
\cite{Zeng:2003:QDW:775152.775211}, Zeng et al. have described a
global planning selection algorithm and a delegation algorithm to
be run when a request to execute an operation is received by the
community. This needs a central entity to run those algorithms.
Such entity plays the same role as the community coordinator or
manager. All these proposals have in common the consideration
of the central entity for community management, which includes the hiring and firing of the community members, while our approach is fully distributed. Moreover, the stability of the community has not been investigated. Another major difference with our work is that the decision making process is only one way and unilateral from the community manager perspective. In our proposal, both participants, namely communities and web services, participate in the decision making process and a joining decision is made only if it is the best option for both players.

@@

>>>>

\textbf{Contributions.} In this chapter, we introduce DDM, a Distributed Decision Making model for community formation that regulates web service agents’ decision making process in terms of cooperating and deciding which group to join and which service to invite for joining. Unlike existing work on community formation, our decision model is extracted from a data model in the form of information obtained from a large number of web services regarding their single and cooperative utilities as well as environmental parameters such as demand, service quality, etc. The generated decision tree improves agents' understanding of the environment and how to select actions that lead towards maximizing their utilities. The advantage of this approach is that the tree, which is initially created from the past data, reflects a comprehensive vision about agents' attitudes in terms of their action selection based on their past experiences. Moreover, the tree is getting continuously updated based on both new received feedback and the outcome of chosen actions. This continuous update makes the approach adapted to any change in the environment.
%The training model deploys a logistic regression algorithm to build a hypothesis function that can thoroughly address the aforementioned research problems.
The decision model provides web services with enough information which helps those services efficiently decide and predict the outcome of their different possible collaborations. This model works in a distributed manner in which services are self-sufficient in their decision making and do not rely on a centralized decision making process. Our findings show that communities of web services can efficiently find the appropriate web service to invite for cooperation as well as allowing a single web service to find the best communities to join. The proposed model can be seen as a recommneder system that suggests beneficial actions for both communities and single services. Communities can consider the decision model and analyze the characteristics of different individual web services and make prudent decisions when inviting a web service to join or accepting a join inquiry initiated from a web service. In general, DDM equips web services with efficient methods for foreseeing how their choices will impact both their short-term and long-term goals; therefore, opting for the best decision available.

@@

We are not going to delve into the engineering details of online web service implementation and its protocols in this thesis. We are interested in web services from a business model perspective. Service providers usually charge end users for services they provide. For example, Google has listed pricing and plans for a wide range of services they provide on their web service console page\footnote{$https://code.google.com/apis/console$}.

@@

In this section, we discuss the parameters that we use in the rest of the chapter. Then, we present the task distribution and revenue model of our distributed web services communities.

@@

With a group of web services having identical or similar functionalities, QoS metrics provide nonfunctional characteristics for optimal candidate selection. Web services quality metrics have been studied and analyzed in various proposals, for instance in in \cite{Ardagna:2007:ASC:1263152.1263531,Menasce:2002:QIW:613357.613758,10.1109/ISSRE.2011.17}. In this chapter, we adopt the most representative QoS properties of those services that highly influence their utility.
%We refer to a typical web service as ws_{i}.

@@

The closest related work \cite{10.1109/TSC.2012.12,
DBLP:conf/IEEEscc/LimTMB12, DBLP:conf/IEEEscc/KhosravifarABT11} and our
previous work \cite{journal-community-formation} regarding the community formation problem have considered a centralized approach where a community manager has complete information of all the web services and their quality
metric and parameters. Those proposals run complex algorithms through all the space of
solutions in order to find the optimal answer. However, in this research work, we
have considered an unexplored and more realistic situation where information is incomplete and a decision
profile is generated based on a smaller set of web services. Our solution helps
communities and web services select actions that lead towards
maximizing their utilities. Therefore, considering the different settings,
we cannot experimentally compare our work with the mentioned related work.

@@

In this direction, recently \cite{DBLP:conf/IEEEscc/LimTMB12,
DBLP:conf/IEEEscc/KhosravifarABT11, 10.1109/TSC.2012.12} proposed
mechanisms to help users and services maximize their gain. A
two-player non-cooperative game between web services and community
master was introduced in
\cite{DBLP:conf/IEEEscc/KhosravifarABT11}. In this game-theoretic
model, the strategies available to a web service when facing a new
community are requesting to join the community, accepting the
master's invitation to join the community, or refusing the
invitation to join. The set of strategies for communities are
inviting the web service or refusing the web service's join
request. Based on their capacity, market share and reputation, the
two players have different sets of utilities over the strategy
profiles of the game. The main limits of this game model are: 1)
its consideration of only three quality parameters, while the
other factors are simply ignored; and 2) the non-consideration of
the web services already residing within the community. The game
is only between the community master and the new web service, and
the inputs from all the other members and their influence on the
master's decision are simply ignored. The consideration of those
inputs and this influence factor is a significant issue as
existing web services can lose utility or payoff because of the
new member, which can result in an unhealthy and unstable group.
The problem comes from the fact that the existing members should
collaborate with the new web services, so probably their
performance as a group can suffer. Existing members may even
deviate and try to join other communities if they are unsatisfied.
Those considerations of forming stable and efficient coalitions
are the main contributions of our thesis.

@@

In \cite{managing-hela-jalel}, Limam and Akaichi have proposed web
service communities with centralized access across distributed web
services. They have proposed a framework for web service
management, query resolution among communities and a query caching
mechanism executed by the manager to improve the performance of
query resolution process among many distributed communities. The
key idea is to cache previous computed results for answering
future queries. Maamar et al. initially in \cite{conf/webist/MaamarLBTS07} and
then comprehensively in \cite{DBLP:journals/ijebr/MaamarSTBB09}
proposed an architecture utilizing \emph{Contract-Net} protocol
for engineering task distribution within communities of web
services. The protocol is centrally executed by the community
manager. This architecture has been further extended in
\cite{CSTintercommunity, conf/IEEEscc/BenharrefSBB11,
conf/IEEEscc/KhosravifarBMMT10, conf/aina/LimTM11}. Two types of
roles have been distinguished for community members: masters and
slaves. Master web services are community managers that lead
communities and are responsible for membership management. They
can invite and convince slave web services to join the community,
and attract new slave web services to their communities by
awarding them better payoff. Moreover, they can eject some slave
members from the community to improve its overall reputation if
these members are misbehaving or cannot provide the promised QoS.
In \cite{Medjahed05adynamic}, Medjahed and Bouguettaya have
developed a community as a ``cluster'' that groups Web services
based on a specific area of interest. All web services in a given
community share the same functionality. These communities are
created by \emph{third party community providers} which use the
\emph{community ontology} as a template and define a set of
operations that all web services within a community should
provide. Using semantic analysis on web service operations, web
services either find and join a community with similar
functionality or create a new operation description for a new
community. The authors have described the concept of
\emph{community agents} associated to \emph{community providers}.
A community agent is responsible, among other things, of the
registration of services with the community. An example of a
community that provides health care services to senior citizens
has been used. In this example, a governmental entity is needed to
check the health care standards used by the members before
authorizing them to be part of the community. Such a central
entity is represented by the community agent. Thus, community
agents are playing the role of community managers. In a close work
\cite{Zeng:2003:QDW:775152.775211}, Zeng et al. have described a
global planning selection algorithm and a delegation algorithm to
be run when a request to execute an operation is received by the
community. This needs a central entity to run those algorithms.
Such entity plays the same role as the community coordinator or
manager. All these proposals have in common the consideration
of the central entity for community management, which includes the hiring and firing of the community members, while our approach is fully distributed. Moreover, the stability of the community has not been investigated. Another major difference with our work is that the decision making process is only one way and unilateral from the community manager perspective. In our contribution, both participants, namely communities and web services, participate in the decision making process and a joining decision is made only if it is the best option for both players.

@@

In next chapter, we analyse the internal community behaviour of web services, and propose a model for competing and cooperating agents within the communities of web services.

### 2015-07-10 thesis/thesis/chap5/chap5.tex
changes

\begin{proof}\footnote{In this paper, we
assume that the common arithmetic and elementary functions, such
as multiplication, division and trigonometric functions can be
computed in time $O(1)$ as they operate on inputs of fixed sizes.}
The function $Rep_{w}^{t}$ is recursive on $t$, but the algorithm
works by storing the last calculated reputation value in a
variable, so it will not be recalculated again at each iteration.
However, the calculation of $reward_w^t$ is needed. Since the
function $reward_w^t$ can be computed in time linear in the number
of tasks (see Equations \ref{repr} and \ref{reward}), which in
turn is linear in the size of the window time, the result follows.
$\Box$
\end{proof}

@@

In Equation \ref{CMPff}, $E^t_w$ is the number of tasks that $w$
expects during the window time $t$, and $\mu_{w, CM}$ is the mean
service fee that is assigned by the master agent to $w$. This
means that a competing service directly obtains this fee from the
master agent. Moreover, the competing service $w$ expects a
cooperation fee ($COF_w^t$) that it gives to its collaborators in
case $w$ needs to cooperate with other services (cooperative
service agents in its collaboration network). In any case, the
competing or cooperating service agent pays a fixed amount of
membership ($\epsilon$) to the master agent, which plays the role
of the community's coordinator. This fee would be taken into
account when a service decides to leave to a cheaper community or
act alone. But to concentrate on the main concerns of this paper,
we skip these small details.\\\\
%
From Equation \ref{CMPff}, the following proposition is
straightforward.

>>>>

\begin{proof}\footnote{In this work, we
assume that the common arithmetic and elementary functions, such
as multiplication, division and trigonometric functions can be
computed in time $O(1)$ as they operate on inputs of fixed sizes.}
The function $Rep_{w}^{t}$ is recursive on $t$, but the algorithm
works by storing the last calculated reputation value in a
variable, so it will not be recalculated again at each iteration.
However, the calculation of $reward_w^t$ is needed. Since the
function $reward_w^t$ can be computed in time linear in the number
of tasks (see Equations \ref{repr} and \ref{reward}), which in
turn is linear in the size of the window time, the result follows.
$\Box$
\end{proof}

@@

In Equation \ref{CMPff}, $E^t_w$ is the number of tasks that $w$
expects during the window time $t$, and $\mu_{w, CM}$ is the mean
service fee that is assigned by the master agent to $w$. This
means that a competing service directly obtains this fee from the
master agent. Moreover, the competing service $w$ expects a
cooperation fee ($COF_w^t$) that it gives to its collaborators in
case $w$ needs to cooperate with other services (cooperative
service agents in its collaboration network). In any case, the
competing or cooperating service agent pays a fixed amount of
membership ($\epsilon$) to the master agent, which plays the role
of the community's coordinator. This fee would be taken into
account when a service decides to leave to a cheaper community or
act alone. But to concentrate on the main concerns of this thesis,
we skip these small details.\\\\
%
From Equation \ref{CMPff}, the following proposition is
straightforward.
